Versatile Robust Clustering of Ad Hoc Cognitive Radio Network

### 2016-09-26 journal-draft/bare_jrnl_comsoc.tex
modified the centralised section

Forming clusters is conducted in the very beginning or periodically according to the dynamics of the CRN.
To form a cluster which is composed by the CR users, there should be channels available for communication within the cluster.
Usually there are multiple unlicensed channels available for all the CR nodes in the cluster, which are referred to in the following of this paper as \textit{common control channels} (CCC).
Out of the available CCCs, there are always one or multiple channels which are used for the payload communication.
%The communication within a cluster is conducted in the spectrum which is available for every member in that cluster.
When one or several cluster members can not use one certain CCC for example because primary user activity has been detected on that channel, it will be excluded from the set of CCCs.
If the channel is being used for payload communication, all the cluster members will switch to another channel in the set of available CCCs to achieve seamless transmission.
As long as there is a CCC available, the corresponding cluster can continue to exchange payload data.
In the context of CRN, as the activity of primary users is unknown to the secondary users, the availability of CCCs of the formed clusters is totally dependent on primary users' activity.
Thus, the clusters with more CCCs are in general better as this anticipates a longer time span that payload can be exchanged.
In this paper the robustness of a cluster against primary users is synonymous with the affluence of CCCs which are possessed by that cluster. 
%In other words, the availability of CCCs of the formed clusters is passive and can not be guaranteed.
%Thus, the number of CCCs in one formed cluster indicates the robustness of it when facing the unpredictable influence from primary users.
%In CRN, one cluster survives the influence of primary users when at least one CCC is available in that cluster.
%As the primary users' operation is assumed to be unknown to the CR users, a cluster formed with more CCCs will survive with higher probability.

%It is easy to see that forming clusters with different neighbours leads to different amount of CCCs in the clusters.
%As a result, how to form the clusters plays an important role on the robustness of clusters in CRN.
%Furthermore, the clustering algorithm also determines the connectivity between several clusters which ultimately determines the robustness of the entire network on the respect of connectivity. 
In the formation process, there is a trade-off between CCCs and the collaborative sensing performance, as less members in a cluster yield in general a higher number of CCCs while this reduces the sensing accuracy~\cite{Consensus_based_clustering12}. 
In addition, there are other reasons that motivate clusters to have more members.
For example, instead of less members, the proper cluster size results in smaller power consumption~\cite{clustering_globecom11, EnergyEfficientClusteringRouting_2015}.
%
%To solely pursue cluster robustness against the primary users' activity, \ie to achieve more common channels within clusters, the ultimately best clustering strategy is ironically that every node constitutes one cluster (called as \textit{singleton cluster} in this paper).
%%In that case, the common channels within cluster are the available channels available at that node's place.
%Apparently this contradicts our motivation of proposing cluster in cognitive radio network to enable cooperative decision making.
Thus the cluster robustness discussed in terms of number of CCCs carries little meaning when the sizes of formed clusters are not given consideration.

@@

The clustering schemes proposed in~\cite{Zhao07, Chen07,Affinity_clustering_09icccn} form clusters which meet the basic condition, having CCC in every cluster.

@@

A clustering scheme (denoted by SOC) which is designed to generate robust clusters against primary users is proposed in~\cite{LIU_TMC11_2}.
SOC involves three phases of distributed executions.
In the first phase, every secondary user forms clusters with some one-hop neighbors, in the second and third phase, each secondary user seeks to either merge other clusters around it, or join one of them.
The metric adopted by every secondary user in the three phases is the product of the the number of CCCs and cluster size (the number of secondary users in the cluster).
%To the best of our knowledge, SOC is the only work emphasizing on the robustness of clustering structure among the related works. 
But this robust clustering scheme has some drawbacks.
Although the adopted metric covers both cluster size and the number of CCCs, the formation of certain clusters may be easily dominated by only one factor, e.g. a node which is able to use many channels will exclude its neighbor and form a cluster by itself.
%emphasis on improving the numbers of common channels within clusters, in order to strengthening robustness of clusters, but the perused metric is not examined or proved to be able to maintain cluster structure under vigorous intensity of primary users’ activities.
%The authors consider the balance between the number of idle common channels within cluster and cluster size and propose an algorithm that increases the number of common channels within clusters. 
%However, this work neglects the issue of connectivity between clusters. 
%One drawback of this scheme is, in order to increase the number of common channels within clusters, the scheme excludes certain CR nodes from the formed clusters, so that isolated nodes have to form clusters themselves. 
In addition, this scheme leads to the high variance of the cluster sizes, which is not desired in certain applications as discussed in~\cite{clustering_globecom11, cluster_EW10}.
%The first two propose a distributed scheme where the metric is the product of cluster size and the number of common control channels.
\cite{mansoor_15_cluster_robust} presents a heuristic method to form clusters, although the authors claim robustness is one goal to achieve, the minimum number of clusters is pursued.
\cite{Li11_ROSS} proposes a distributed clustering scheme under the game theoretic framework. 
Compared with the clustering schemes introduced so far, the clusters are formed in shorter time and posses more CCCs within and among clusters.
But this scheme doesn't consider the the factor of size in the algorithm design.

@@

\section{System Model}
\label{sec:model}
%Let us consider a two dimensional area where primary and secondary users coexist together.
%They share the licensed channels according to the spectrum overlay model, where secondary transmitters are only allowed to transmit when the prifcan exchange its spectrum sensing resulmary users are detected as being idle, and they should vacate the working channel when the presence of primary user is detected.
%The set of primary users and secondary users are presented by $\mathcal{P}$ and $\mathcal{N}$ separately, there are $|\mathcal{P}| = P$ and $|\mathcal{N}| = N$.
%Cognitive radio ad hoc network is constituted by all the secondary users in $\mathcal{N}$.
%The collection of non-overlapping licensed frequency bands is denoted as $\mathcal{F}$ with $|\mathcal{F}| =F$, which is shared by the the primary and secondary users.
%We assume that primary users have a relatively low variation in activity (periods of activity and inactivity in the range of seconds or minutes).
%%
%Primary users access the allocated channels in $\mathcal{F}$ according to its need without sending any explicit notification to secondary users.

We consider a set of cognitive radio users $\mathcal{N}$ and a set of primary users distributed over a given area.
%We denote the set of these nodes as $V$ which is partitioned into sets of primary users $V_P$ and secondary users $V_N$, i.e., $V~=~V_P~\dot\cup~V_N$. 
%These users share a number of non-overlapping licensed channels according to the spectrum overlay model. 
The CR users use a set of licensed channels opportunistically, which is denoted as $\mathcal{K}$. 
The CR users are allowed to transmit on channel $k \in \mathcal{K}$ only if no primary user is detected on channel $k$. 
%The available channels sensed on secondary user $i$ is denoted by $V_i$ and there is $\vert V_i \vert \leq F$. % indicating the total number of 
%%Each secondary user notifies its neighbours of the sensing result about the available channels.
CR users conduct spectrum sensing independently and sequentially on all licensed channels.
%The sensing duration and frequency on one channel is a research topic~\cite{sensing_survey_2009}, and 
We assume that every node can detect the presence of a primary user on each channel with certain accuracy. \footnote{The spectrum availability can be validated with a certain probability of detection. Spectrum sensing/validation is out of the scope of this paper.}
We adopt the unit disk model~\cite{unitDiskModel} for the transmission of both primary and CR users.
%Both primary users and CR users have fixed transmission ranges respectively, and all the channels are regarded to be identical in terms of signal propagation.
If a CR node locates within the transmission range of an active primary user, that CR node is not allowed to use the channel which is being used by that primary user.
%In our system model, we assume that a secondary node $n$ is able to detect for each channel $c$ whether any primary user utilizes $c$.
As the result of spectrum sensing, $K_i \subseteq \mathcal{K}$ denotes the set of licensed channels which can be used by $i$.
%
%One dedicated control channel is assumed to be available for all the secondary nodes to exchange control messages with neighbours in the process of cluster formation.
%Note that the assumption of dedicated control channel is to simplify the discussion so that we can focus on the kernel of this paper, the robustness of clusters.
%Actually, the control messages involved in the clustering process can be conveyed on available licensed channels through rendezvous process by channel hopping~\cite{channelHopping_Rendezvous_2014, Gu_distributed_rendezvous_2014}. 
%Over the control channel, secondary users exchange their spectrum sensing results $V_{i}$ with one hop neighbours. 
%Secondary users have the same transmission range on both licensed and control channel, and the communication links are reciprocal.
%

We assume that in addition to the licensed channels, there is one dedicated control channel.
This control channel could be one of the ISM bands or other reserved spectrum which is exclusively used for transmitting control messages.
\footnote{Actually, the control messages involved in the clustering process can also be transmitted on the available licensed channels through a rendezvous process by channel hopping~\cite{channelHopping_Rendezvous_2014, Gu_distributed_rendezvous_2014}, i.e., two neighbouring nodes establish communication on the same channel.}
Over the control channel, a secondary user $i$ can exchange its spectrum sensing result $K_i$ to all its one hop neighbors $\text{Nb}(i)$, where $\text{Nb}(i)$ is the neighbors of $i$, and is simply defined as the set of CR nodes located within the transmission range of $i$. 
%It is available for any secondary node $i$ to exchange control messages with any other node in its neighborhood $\text{Nb}(i)$ during the cluster formation phase.
In the following, we refer to the word \textit{channel} in general as licensed channel, unless we mention the control channel explicitly.

%Primary users also have fixed transmission range on the licensed channels\footnote{This assumption is made to simplify the discussion, and doesn't affect the effectiveness of the proposed scheme}.

%As to a pair of secondary users, when the Euclidean distance between them is smaller than secondary users' transmission range, we assume the pair is able to communicate on both control channel and licensed channel, and the both are considered to be neighbours of each other. DEFINE DISTANCE! \textcolor{brown}{distance, done}
%
If a secondary user $i$ is not in the transmission range of an active primary user $p$, $i$ can certainly not detect the presence of $p$.
As the transmission range of primary users is limited and secondary users have different locations, different secondary users may have different views of the spectrum availability, i.e., for any $i, j \in \mathcal{N}$, $K_i = K_{j}$ does not necessarily hold.	
%
%Neighbourhood establishment and maintenance with control channel and according to a neighborhood discovery protocol which is out of scope of our work.
%While primary users are assumed to be fixed, secondary users can be mobile.
%Validation refers to the process of ensuring that no primary transmission is actually taking place on the respective channel.
%All $N$ secondary users constitute an ad-hoc network in which data can be transmitted from one certain node to any other node should available c.hannels are available on the source node, destination node and other intermediate nodes.
%Communication on licensed channels is possible only when nodes $i, j$ are both located in each other's transmission range and both share a validated licensed channel.
%As the assumed $0/1$ state of connectivity is solely based on the Euclidean distance between secondary users,
%CRN can be represented by a graph $\mathcal{G}(I,\mathcal{E})$, where $\mathcal{E}=\lbrace(i,j,v) \vert i, j \in \mathcal{N} \wedge v\in V_i \wedge v\in V_j \rbrace$ is wireless link between any secondary node $i$ and its neighbour $j$ with licensed channel $v$.
%Due to relatively low primary user dynamics, time index is omitted here.
%%In order to perform clustering, CR users first need to establish their neighborhood.
%For secondary node $i$, its neighborhood $Nb_i$ consists of all the secondary users locating within its transmission range, regardless whether common licensed channels exist or not. 
%% and have at least one common channel with node $i$ each, i.e. $ j\in Nb_i \Rightarrow V_i\cap V_j\neq \emptyset$.
%%The clustering phase is initialized during which any control message is again conveyed by the control channel.
%In the rest of this paper, \textit{channel} only refers to the licensed common control channel unless the dedicated control channel is particularly mentioned.
Then, a cognitive radio network can be represented as a graph $G = (\mathcal{N}, E)$, where $E \subseteq \mathcal{N} \times \mathcal{N}$ such that $\{i, j\} \in E$  if, and only if, there exists a channel $k \in \mathcal{K}$ with $k \in K_{i} \cap K_{i'}$. 
%Note that the channel availability is that when the clustering process is performed. 
%For the rest of this paper the word channel is referred to licensed channel, if the control channel is not explicitly mentioned.

\section{Robust Clustering Problem in CRN}
\label{problem}
In this section, we present the definition of cluster in the context of CRNs, define the clustering problem and discuss its complexity.

\subsection{Cluster in CRN}
\label{def_cluster}
%We give the description of cluster in the context of CRN, all the clusters discussed in this paper comply with it.
%A cluster $C$ is composed with one cluster head and cluster members, which satisfies the following conditions:
%\begin{itemize}
%\item Cluster head $H_C$ is able to communicate with any cluster member \textit{directly}, \ie for any cluster member $i\in C$, there is $i\in \texttt{Nb}_{H_C}$.
%\item There exists as least one common control channel in the cluster, \ie $\cap_{i\in C} V_i \neq \emptyset$.
%\end{itemize}
A cluster $C$ is a set of secondary nodes consisting of a cluster head $h_C$ and a number of cluster members.
The cluster head is able to communicate with any cluster member directly.
In other terms, for any cluster member $i \in C$, $i \in \text{Nb} (h_C) $ holds.

%A cluster head coordinates the activities of cluster members, \ie it notifies all the members to vacate a channel if a primary user's presence is detected by one cluster member. Further it also notifies the members to use a different common control channel for payload communication. %(\textcolor{red}{(Not clear what the meaning of payload communication is. Why not omitting payload here?)} \textcolor{brown}{payload communication happens when the clusters are formed, and start to support network functionalities, e.g. routing. Thus I think it is necessary to leave it here.}

In this paper, sometimes we need to mention a cluster with the information of its cluster head \ie a cluster is denoted as $C(i)$ where $i$ is its cluster head.
Cluster size of $C(i)$ is written as $|C(i)|$.
%, for the sake of concision, $C_i$ is also named as cluster $i$ in the following part of this paper.
$ K(C) = \cap_{i\in C} K_i$, $K(C)$ denotes the set of common control channels in cluster $C$.
Clustering is performed periodically, as secondary users are mobile and the channel availability on secondary users are changing as primary users change their operation state.%, or the most current clusters can not maintain due to lack of CCCs when primary users' operation is intense.

@@

\begin{table}[h!]
\caption{Notations}
\label{tab1}
\centering
\begin{tabular}{llr}
\toprule
%\multicolumn{2}{c}{Item} \\
%\cmidrule(r){1-2}
Symbol & Description \\
\midrule
$\mathcal{N}$  & collection of CR users in a CRN\\
$N$ & number of CR users in a CRN, $N=|\mathcal{N}|$\\
$\mathcal{K}$	& set of licensed channels\\
$k(i)$ & the working channel of user $i$\\
$\text{Nb}(i)$ & the neighborhood of CR node $i$    \\
$C(i)$ & a cluster whose cluster head is $i$  \\
%$k_i$   & set of available channels at CR node $i$  \\
$K_i$   & the set of available channels at CR node $i$  \\
$K(C(i))$   & the set of available CCCs of cluster $C(i)$ \\
%$h(C)$ & cluster head of a cluster $C$\\
$h_C$ & the cluster head of a cluster C\\
$\text{CH}$ & cluster head\\
%$\text{CH}$ & cluster head\\
$\delta$ & the cluster size which is preferred\\
$S_i$ & a set of claiming clusters, each of which includes \\
& debatable node $i$ after phase I\\
$d_i$  & individual connectivity degree of CR node $i$\\
$g_i$  & neighborhood connectivity degree of CR node $i$\\
$C_i$  & the $i$th potential cluster (only used in Sec.~\ref{centralized_opt})\\
\bottomrule
\end{tabular}
\end{table}

\subsection{Robust Clustering Problem}
\label{centralized_scheme}

%paper given by james is useful which provides a survey from the perspective of wsn!

The clustering scheme we propose in the paper aims to form clusters with preferred sizes, meanwhile maximize the average number of CCCs per cluster.
%In the following, we refer this problem as \textit{centralized clustering}. 
%We have mentioned that the sizes of the formed clusters should be given consideration, which is import for the discussion on cluster robustness, \ie 
when perusing the cluster robustness without specifying the requirement on cluster sizes, a clustering method will be inclined to form small clusters.
In that case, although more CCCs are obtained per cluster, the benefits brought in by the collective of the cluster members are compromised.
%Except for cooperative sensing, clusters need to conduct some other functionalities.
Large clusters are not preferred in some scenarios neither, \eg for the CRN composed with resource limited users, managing the cluster members is a substantial burden. 
Hence, the cluster size should fall in a desired range~\cite{Chen04clusteringalgorithms, capacity_cluster_06}.
In the following, we present the formal definition of robust clustering problem.

@@

Given a cognitive radio network $\mathcal{N}$ where $|\mathcal{N}|=N$, secondary users are indexed from 1 to $N$.
A cluster $C$ can be composed with one or multiple secondary users when $|K(C)|>0$, and any two cluster members can communicate either directly, or through other members belonging to the same cluster.
We use $f(C)$ to denote the number of CCCs of a cluster $C$, and give it a new definition as follows, the number of common control channels is $f(C) = |K(C)|$ if $|C|>1$, and $f(C)=0$ when $|C|=1$.
The size of cluster $C$ is not bigger than a positive integer $k$.
The collection of such clusters $\mathcal{S}=\{C_1, C_2,\ldots,C_{|\mathcal{S}|}\}$, satisfies the following condition: $\bigcup_{1\leq i \leq |\mathcal{S}|} C_i = N$ and $1\leq i \leq |\mathcal{S}|$.

%, which distinguishes the centralized clustering problem discussed in this paper.

The robust clustering problem is to find a subcollection $\mathcal{S}' \subseteq \mathcal{S}$, so that $\sum_{C\in \mathcal{S}'} f(C)$ is maximized, and $\mathcal{S}'$ satisfies the following conditions, $\bigcup_{C_j\in \mathcal{S}'} C_j = \mathcal{N}$ and $C_{j'}\cap C_j =\emptyset$ where $C_{j'}, C_j\in \mathcal{S}'$ and $j'\neq j$.
% when $|C|>1$, and $f(\cdot)=0$ when $|C|=1$.
%As to $f(\cdot):z\rightarrow z$, under this problem setting, if only number of common channels are given consideration, only singleton clusters will be preferred, which contradicts to our goal of clustering CR nodes together, thus we choose $f(\cdot)= |K_{C}|\cdot |C|$, the product of number of common channels and cluster size.
\end{mydef}

>>>>

Forming clusters is conducted in the very beginning or periodically according to the dynamics of the CRN.
To form a cluster which is composed by the CR users, there should be channels available for communication within the cluster.
Usually there are multiple unlicensed channels available for all the CR nodes in the cluster, which are referred to in the following of this paper as \textit{common licensed channels} (or common channels for short, which is abbreviated as CC).
Out of the available CCs, there are always one or multiple channels which are used for the payload communication.
%The communication within a cluster is conducted in the spectrum which is available for every member in that cluster.
When one or several cluster members can not use one certain CC for example because primary user activity has been detected on that channel, it will be excluded from the set of CCs.
If the channel is being used for payload communication, all the cluster members will switch to another channel in the set of available CCs to achieve seamless transmission.
As long as there is a CC available, the corresponding cluster can continue to exchange payload data.
In the context of CRN, as the activity of primary users is unknown to the secondary users, the availability of CCs of the formed clusters is totally dependent on primary users' activity.
Thus, the clusters with more CCs are in general better as this anticipates a longer time span that payload can be exchanged.
In this paper the robustness of a cluster against primary users is synonymous with the affluence of CCs which are possessed by that cluster. 
%In other words, the availability of CCs of the formed clusters is passive and can not be guaranteed.
%Thus, the number of CCs in one formed cluster indicates the robustness of it when facing the unpredictable influence from primary users.
%In CRN, one cluster survives the influence of primary users when at least one CC is available in that cluster.
%As the primary users' operation is assumed to be unknown to the CR users, a cluster formed with more CCs will survive with higher probability.

%It is easy to see that forming clusters with different neighbours results in different number of CCs in the clusters.
%As a result, how to form the clusters plays an important role on the robustness of clusters in CRN.
%Furthermore, the clustering algorithm also determines the connectivity between several clusters which ultimately determines the robustness of the entire network on the respect of connectivity. 
In the formation process, there is a trade-off between CCs and the collaborative sensing performance, as less members in a cluster yield in general a higher number of CCs while this reduces the sensing accuracy~\cite{Consensus_based_clustering12}. 
In addition, there are other reasons that motivate clusters to have more members.
For example, instead of less members, the proper cluster size results in smaller power consumption~\cite{clustering_globecom11, EnergyEfficientClusteringRouting_2015}.
%
%To solely pursue cluster robustness against the primary users' activity, \ie to achieve more common channels within clusters, the ultimately best clustering strategy is ironically that every node constitutes one cluster (called as \textit{singleton cluster} in this paper).
%%In that case, the common channels within cluster are the available channels available at that node's place.
%Apparently this contradicts our motivation of proposing cluster in cognitive radio network to enable cooperative decision making.
Thus the cluster robustness discussed in terms of number of CCs carries little meaning when the sizes of formed clusters are not given consideration.

@@

The clustering schemes proposed in~\cite{Zhao07, Chen07,Affinity_clustering_09icccn} form clusters which meet the basic condition, having CC in every cluster.

@@

A clustering scheme (denoted by SOC) which is designed to generate robust clusters against primary users is proposed in~\cite{LIU_TMC11_2}.
SOC involves three phases of distributed executions.
In the first phase, every secondary user forms clusters with some one-hop neighbors, in the second and third phase, each secondary user seeks to either merge other clusters around it, or join one of them.
The metric adopted by every secondary user in the three phases is the product of the the number of CCs and cluster size (the number of secondary users in the cluster).
%To the best of our knowledge, SOC is the only work emphasizing on the robustness of clustering structure among the related works. 
But this robust clustering scheme has some drawbacks.
Although the adopted metric covers both cluster size and the number of CCs, the formation of certain clusters may be easily dominated by only one factor, e.g. a node which is able to use many channels will exclude its neighbor and form a cluster by itself.
%emphasis on improving the numbers of common channels within clusters, in order to strengthening robustness of clusters, but the perused metric is not examined or proved to be able to maintain cluster structure under vigorous intensity of primary users’ activities.
%The authors consider the balance between the number of idle common channels within cluster and cluster size and propose an algorithm that increases the number of common channels within clusters. 
%However, this work neglects the issue of connectivity between clusters. 
%One drawback of this scheme is, in order to increase the number of common channels within clusters, the scheme excludes certain CR nodes from the formed clusters, so that isolated nodes have to form clusters themselves. 
In addition, this scheme leads to the high variance of the cluster sizes, which is not desired in certain applications as discussed in~\cite{clustering_globecom11, cluster_EW10}.
%The first two propose a distributed scheme where the metric is the product of cluster size and the number of common channels.
\cite{mansoor_15_cluster_robust} presents a heuristic method to form clusters, although the authors claim robustness is one goal to achieve, the minimum number of clusters is pursued.
\cite{Li11_ROSS} proposes a distributed clustering scheme under the game theoretic framework. 
Compared with the clustering schemes introduced so far, the clusters are formed in shorter time and posses more CCs within and among clusters.
But this scheme doesn't consider the the factor of size in the algorithm design.

@@

\section{System Model and Problem Formulation}
\label{sec:model}
%Let us consider a two dimensional area where primary and secondary users coexist together.
%They share the licensed channels according to the spectrum overlay model, where secondary transmitters are only allowed to transmit when the prifcan exchange its spectrum sensing resulmary users are detected as being idle, and they should vacate the working channel when the presence of primary user is detected.
%The set of primary users and secondary users are presented by $\mathcal{P}$ and $\mathcal{N}$ separately, there are $|\mathcal{P}| = P$ and $|\mathcal{N}| = N$.
%Cognitive radio ad hoc network is constituted by all the secondary users in $\mathcal{N}$.
%The collection of non-overlapping licensed frequency bands is denoted as $\mathcal{F}$ with $|\mathcal{F}| =F$, which is shared by the the primary and secondary users.
%We assume that primary users have a relatively low variation in activity (periods of activity and inactivity in the range of seconds or minutes).
%%
%Primary users access the allocated channels in $\mathcal{F}$ according to its need without sending any explicit notification to secondary users.

We consider a set of cognitive radio users $\mathcal{N}$ and a set of primary users distributed over a given area.
%We denote the set of these nodes as $V$ which is partitioned into sets of primary users $V_P$ and secondary users $V_N$, i.e., $V~=~V_P~\dot\cup~V_N$. 
%These users share a number of non-overlapping licensed channels according to the spectrum overlay model. 
The CR users use a set of licensed channels opportunistically, which is denoted as $\mathcal{K}$. 
The CR users are allowed to transmit on channel $k \in \mathcal{K}$ only if no primary user is detected on channel $k$. 
%The available channels sensed on secondary user $i$ is denoted by $V_i$ and there is $\vert V_i \vert \leq F$. % indicating the total number of 
%%Each secondary user notifies its neighbours of the sensing result about the available channels.
CR users conduct spectrum sensing independently and sequentially on all licensed channels.
%The sensing duration and frequency on one channel is a research topic~\cite{sensing_survey_2009}, and 
We assume that every node can detect the presence of an active primary user on each channel with certain accuracy. \footnote{The spectrum availability can be validated with a certain probability of detection. Spectrum sensing/validation is out of the scope of this paper.}
We adopt the unit disk model~\cite{unitDiskModel} for the transmission of both primary and CR users.
%Both primary users and CR users have fixed transmission ranges respectively, and all the channels are regarded to be identical in terms of signal propagation.
If a CR node locates within the transmission range of an active primary user, that CR node is not allowed to use the channel which is being used by that primary user.
%In our system model, we assume that a secondary node $n$ is able to detect for each channel $c$ whether any primary user utilizes $c$.
As the result of spectrum sensing, $K_i \subseteq \mathcal{K}$ denotes the set of licensed channels which can be used by $i$.
%
%One dedicated control channel is assumed to be available for all the secondary nodes to exchange control messages with neighbours in the process of cluster formation.
%Note that the assumption of dedicated control channel is to simplify the discussion so that we can focus on the kernel of this paper, the robustness of clusters.
%Actually, the control messages involved in the clustering process can be conveyed on available licensed channels through rendezvous process by channel hopping~\cite{channelHopping_Rendezvous_2014, Gu_distributed_rendezvous_2014}. 
%Over the control channel, secondary users exchange their spectrum sensing results $V_{i}$ with one hop neighbours. 
%Secondary users have the same transmission range on both licensed and control channel, and the communication links are reciprocal.
%

We assume there is one dedicated control channel which is used to exchange signaling messages during clustering process.
This control channel could be one of the ISM bands or other reserved spectrum which is exclusively used for transmitting control messages.
\footnote{Actually, the control messages involved in the clustering process can also be transmitted on the available licensed channels through a rendezvous process by channel hopping~\cite{channelHopping_Rendezvous_2014, Gu_distributed_rendezvous_2014}, i.e., two neighbouring nodes establish communication on the same channel.}
Over the control channel, a secondary user $i$ can exchange its spectrum sensing result $K_i$ to all its one hop neighbors $\text{Nb}(i)$, where $\text{Nb}(i)$ is the neighbors of $i$, and is simply defined as the set of CR nodes located within the transmission range of $i$. 
%It is available for any secondary node $i$ to exchange control messages with any other node in its neighborhood $\text{Nb}(i)$ during the cluster formation phase.
In the following, we refer to the word \textit{channel} in general as licensed channel, unless we mention the control channel explicitly. 
%
%
If a secondary user $i$ is not in the transmission range of an active primary user $p$, $i$ can certainly not detect the presence of $p$.
As the transmission range of primary users is limited and secondary users have different locations, different secondary users may have different views of the spectrum availability, i.e., for any $i, j \in \mathcal{N}$, $K_i = K_{j}$ does not necessarily hold.	
%
%Neighbourhood establishment and maintenance with control channel and according to a neighborhood discovery protocol which is out of scope of our work.
%While primary users are assumed to be fixed, secondary users can be mobile.
%Validation refers to the process of ensuring that no primary transmission is actually taking place on the respective channel.
%All $N$ secondary users constitute an ad-hoc network in which data can be transmitted from one certain node to any other node should available c.hannels are available on the source node, destination node and other intermediate nodes.
%Communication on licensed channels is possible only when nodes $i, j$ are both located in each other's transmission range and both share a validated licensed channel.
%As the assumed $0/1$ state of connectivity is solely based on the Euclidean distance between secondary users,
%CRN can be represented by a graph $\mathcal{G}(I,\mathcal{E})$, where $\mathcal{E}=\lbrace(i,j,v) \vert i, j \in \mathcal{N} \wedge v\in V_i \wedge v\in V_j \rbrace$ is wireless link between any secondary node $i$ and its neighbour $j$ with licensed channel $v$.
%Due to relatively low primary user dynamics, time index is omitted here.
%%In order to perform clustering, CR users first need to establish their neighborhood.
%For secondary node $i$, its neighborhood $Nb_i$ consists of all the secondary users locating within its transmission range, regardless whether common licensed channels exist or not. 
%% and have at least one common channel with node $i$ each, i.e. $ j\in Nb_i \Rightarrow V_i\cap V_j\neq \emptyset$.
%%The clustering phase is initialized during which any control message is again conveyed by the control channel.
%In the rest of this paper, \textit{channel} only refers to the licensed common channel unless the dedicated control channel is particularly mentioned.
Then, a cognitive radio network can be represented as a graph $G = (\mathcal{N}, E)$, where $E \subseteq \mathcal{N} \times \mathcal{N}$ such that $\{i, j\} \in E$  if, and only if, there exists a channel $k \in \mathcal{K}$ with $k \in K_{i} \cap K_{i'}$. 
%Note that the channel availability is that when the clustering process is performed. 
%For the rest of this paper the word channel is referred to licensed channel, if the control channel is not explicitly mentioned.

%In this subsection, we present the definition of cluster in the context of CRNs, define the clustering problem and discuss its complexity.
%We give the description of cluster in the context of CRN, all the clusters discussed in this paper comply with it.
%A cluster $C$ is composed with one cluster head and cluster members, which satisfies the following conditions:
%\begin{itemize}
%\item Cluster head $H_C$ is able to communicate with any cluster member \textit{directly}, \ie for any cluster member $i\in C$, there is $i\in \texttt{Nb}_{H_C}$.
%\item There exists as least one common channel in the cluster, \ie $\cap_{i\in C} V_i \neq \emptyset$.
%\end{itemize}
A cluster $C$ is a set of secondary nodes, which consists a cluster head $h_C$ and a number of cluster members.
The cluster head is able to communicate with any cluster member directly.
In other terms, for any cluster member $i \in C$, $i \in \text{Nb} (h_C) $ holds.
%
%A cluster head coordinates the activities of cluster members, \ie it notifies all the members to vacate a channel if a primary user's presence is detected by one cluster member. Further it also notifies the members to use a different common channel for payload communication. %(\textcolor{red}{(Not clear what the meaning of payload communication is. Why not omitting payload here?)} \textcolor{brown}{payload communication happens when the clusters are formed, and start to support network functionalities, e.g. routing. Thus I think it is necessary to leave it here.}
%
%
In this paper, sometimes we need to mention a cluster with the information of its cluster head \ie a cluster is denoted as $C(i)$ where $i$ is its cluster head.
Cluster size of $C(i)$ is written as $|C(i)|$.
$ K(C) = \cap_{i\in C} K_i$, $K(C)$ denotes the set of common licensed channels in cluster $C$.
Clustering is performed periodically, as secondary users are mobile and the channel availability on secondary users are changing as primary users change their operation state.%, or the most current clusters can not maintain due to lack of CCs when primary users' operation is intense.

The notations used in the system model and the following problem description are listed in the Table~\ref{tab1}.
\begin{table}[h!]
\caption{Notations}
\label{tab1}
\centering
\begin{tabular}{llr}
\toprule
%\multicolumn{2}{c}{Item} \\
%\cmidrule(r){1-2}
Symbol & Description \\
\midrule
$\mathcal{N}$  & collection of CR users in a CRN\\
$N$ & number of CR users in a CRN, $N=|\mathcal{N}|$\\
$\mathcal{K}$	& set of licensed channels\\
$k(i)$ & the working channel of user $i$\\
$\text{Nb}(i)$ & the neighborhood of CR node $i$    \\
$C(i)$ & a cluster whose cluster head is $i$  \\
%$k_i$   & set of available channels at CR node $i$  \\
$K_i$   & the set of available channels at CR node $i$  \\
$K(C(i))$   & the set of available CCs of cluster $C(i)$ \\
%$h(C)$ & cluster head of a cluster $C$\\
$h_C$ & the cluster head of a cluster C\\
$\text{CH}$ & cluster head\\
%$\text{CH}$ & cluster head\\
$\delta$ & the cluster size which is preferred\\
$S_i$ & a set of claiming clusters, each of which includes \\
& debatable node $i$ after phase I\\
$d_i$  & individual connectivity degree of CR node $i$\\
$g_i$  & neighborhood connectivity degree of CR node $i$\\
$f(C)$ & the number of CCs of a cluster $C$, which is used \\
& in the problem description\\
$C_i$  & the $i$th potential cluster in a collection of clusters \\
\bottomrule
\end{tabular}
\end{table}

@@

\subsection{Robust Clustering Problem in CRN}
\label{problem}
%paper given by james is useful which provides a survey from the perspective of wsn!
As introduced in the section of introduction, in order to be robust against primary users' activity, the formed clusters should have more CCs to expect longer time expectancy.
Meanwhile, the sizes of the formed clusters should not diverge from the desired size too much, and there are two reasons.
First, the generation of the \textit{singleton clusters}, \ie the cluster which has only one CR node, should be avoided, because it contradicts the motivation of forming clusters thus the benefits brought in by the collective of the cluster members are compromised.
Second, Large clusters are not preferred in some scenarios neither, \eg for the CRN composed with resource limited users, managing the cluster members is a substantial burden. 
Hence, the cluster size should fall in a desired range according to different application scenarios~\cite{Chen04clusteringalgorithms, capacity_cluster_06}.
Considering the above mentioned requirements on the formed clusters, we will present the definition of robust clustering problem as follows.
%The clustering scheme we propose in the paper aims to form clusters with preferred sizes, meanwhile maximize the average number of CCs per cluster.
%We have mentioned that the sizes of the formed clusters should be given consideration, which is import for the discussion on cluster robustness, \ie 
%When perusing the cluster robustness without specifying the requirement on cluster sizes, a clustering method will be inclined to form small clusters.
%In that case, although more CCs are obtained per cluster, the benefits brought in by the collective of the cluster members are compromised.
%Except for cooperative sensing, clusters need to conduct some other functionalities.

@@

Given a cognitive radio network $\mathcal{N}$ where $|\mathcal{N}|=N$,  the secondary users are indexed from 1 to $N$.
A cluster $C$ is formed when first, it is composed with one or multiple secondary users and there is $|K(C)|>0$, second, any two cluster members can communicate either directly, or through other members which belong to the same cluster.
The collection of all the possible clusters in $\mathcal{N}$ is denoted as $\mathcal{S}$, there is $\mathcal{S}=\{C_1, C_2,\ldots,C_{|\mathcal{S}|}\}$ where $C_i$ is the $i$ th cluster in $\mathcal{S}$\footnote{The sequence of the clusters can be decided in any convenient way \ie the sequence of identifying them.}.
%the clustering solution will be composed by the elements of $\mathcal{S}$. 
As to $\mathcal{S}$, there is $\bigcup_{1\leq i \leq |\mathcal{S}|} C_i = N$.

To formulate the problem, we propose a new definition for the number of CCs in a cluster.
We use $f(C_i)$ to denote the number of CCs of a cluster $C_i\in \mathcal{S}$, which is as follows,
%and the number of common licensed channels is $f(C_i) = |K(C_i)|$ if $|C_i|>1$, and $f(C_i)=0$ when $|C_i|=1$.
$$
f(C_i) = \left\{ \begin{array}{rl}
|K(C_i)| 	&\mbox{$|C_i|>1$} \\
0 				&\mbox{$|C_i|=1$} \\
\end{array} \right.
$$

%The size of cluster $C$ is not bigger than a positive integer $k$.
%Then the sum of the number of CCs of the formed clusters is $\sum_{C\in \mathcal{S}'} f(C)$
